---
title: "Data processing sources and notes ZMB"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
#bibliography: ../../Common/Global-to-local-GLOBIOM-ZMB.bib
---

```{r setup, include=FALSE}
library(rprojroot)
root <- find_root(is_rstudio_project)

library(knitr)
figPath <- 
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"FigTabMap/out-"))

library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
source(file.path(root, "Code/ZMB/Set_country.R"))
source(file.path(root, paste0("Code/", iso3c_sel, "/report/maps.r")))
#source(file.path(root, paste0("Code/", iso3c_sel, "/report/tabs.r")))
#source(file.path(root, "Code/MWI/Agricultural_statistics/Analyse_land_MWI.R"))
```

# Data sources
@AgroMaps2017 presents data for yield, area and production at adm1 and adm2 level.Data for adm2 data is only available for 1992.

@CropSTATS2017 


# To check
- Suitability maps as presented in IFPRI MWI report, 2016
- Land use map 1991 with more detailed location of various crops. Also check FAO 1990 and 2010 land cover maps

# Maps

\blandscape
```{r adm_gmia_pop, fig.show='hold',fig.align='center'}
fig_adm
fig_pop
fig_gmia
```
\elandscape


```{r land_cover}
#fig_land_cover
```

```{r household_survey}
#fig_LSMS
```


## Agricultural statistics


\blandscape
```{r area_rank}
#pander(tab_area_rank_FAOSTAT, caption = "Area and area share per crop using FAOSTAT (2000)")
```

\pagebreak
```{r tab_area_rank_as}
#pander(tab_area_rank_as, caption = "Area and area share per crop using national agricultural statistics (2007)")
```

\pagebreak
```{r tab_area_rank_ma}
#pander(tab_area_rank_as, caption = "Area and area share per crop using Agro-maps (2000)")
```

\pagebreak
```{r fig_area_compare}
#fig_area_compare
```

\pagebreak
```{r fig_prod_compare}
#fig_prod_compare
```

\pagebreak
```{r fig_yld_compare}
#fig_yld_compare
```

\pagebreak
```{r fig_heads_compare}
#fig_heads_compare
```

\pagebreak
\footnotesize
```{r tab_area_share_within_reg_as}
#pander(tab_area_share_within_reg_as, caption = "Area share per crop within adm using national agricultural statistics")
```
\normalsize

\pagebreak
```{r fig_area_share}
#fig_area_share
```

\pagebreak
```{r fig_yld_outlier}
#fig_yld_outlier
```

\pagebreak
```{r fig_yld_winsor}
#fig_yld_winsor
```

\pagebreak
```{r fig_yld_winsor_2000}
#fig_yld_winsor_2000
```

\elandscape


# Analysis of land cover and land use

```{r adm_land_cover}
# land cover by adm
```

```{r lu_lc_comparison}
# Comparison between land use from ag_stat and land cover
```

```{r MWI_as2FLC}
# mapping as 2 FAO
```



# References


