---
title: "Data processing sources and notes MWI"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output: pdf_document
bibliography: ../../Common/Global-to-local-GLOBIOM-MWI.bib
---

```{r setup, include=FALSE}
library(rprojroot)
root <- find_root(is_rstudio_project)

library(knitr)
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"FigTabMap/out-"),
  dev = "CairoPNG",
  dev.args = list(CairoPNG = list(bg = "transparent"))
  )

library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

source(file.path(root, "Code/MWI/Agricultural_statistics/Analyse_ag_stat_MWI.R"))
```

# Data sources
@NSO2010 also referred to as NACAL provides the result of a national agricultural census that was undertaken for the year 2006/2007. It provides data at national and district level. 

@NationalStatisticalOffice2012 presents basic information about the hird Integrated Household Survey (IHS3). The data of the survey can be downloaded from the LSMS World Bank website: http://econ.worldbank.org/WBSITE/EXTERNAL/EXTDEC/EXTRESEARCH/EXTLSMS/0,,contentMDK:23635577~pagePK:64168445~piPK:64168309~theSitePK:3358997,00.html. 

@RCMRD2012 describes the procedure to prepare the 2000 and 2010 land cover maps for Malawi.


## Agricultural statistics
- We only have agricultural statistics data at the adm2 GAUL level for 2007. The following procedure was used to project the data to 2000, the base year of GLOBIOM:
  + We created a moving average of FAOSTAT yield series using a window of five years to smooth annual spikes.
  + We calculate the growth in yield between 2001 and 2007.
  + We scale the yield data from national statistics using the national growth rates for each region.
- The tables on production, area and yield in @NSO2010 do not present information on sugar cane cropping. Nonetheless, on p. 264 information is presented that sugar cane is grown in the country  
- Donkeys and rabbits are not matched to a GLOBIOM category
- Check matching of MWI crops to GLOBIOM
- Check land use class aggregation to GLOBIOM
- Data on area om @NSO2010 is for pure stand crop. This might explain why total country values from FAOSTAT are much higher.
- According to FAOSTAT there are no sweet potatoes in MWI, while according to @NSO2010 these form the largest share of potatoes. Probably the FAOSTAT sums both types and classifies them as potatoes


\blandscape
```{r area_rank}
pander(tab_area_rank_FAOSTAT, caption = "Area and area share per crop using FAOSTAT")
```

\pagebreak
```{r tab_area_rank_nat}
pander(tab_area_rank_nat, caption = "Area and area share per crop using national agricultural statistics")
```

\pagebreak
```{r fig_area_compare}
fig_area_compare
```

\pagebreak
```{r fig_prod_compare}
fig_prod_compare
```

\pagebreak
```{r fig_yld_compare}
fig_yld_compare
```

\pagebreak
```{r fig_heads_compare}
fig_heads_compare
```

\pagebreak
\footnotesize
```{r tab_area_share_within_reg}
pander(tab_area_share_within_reg, caption = "Area share per crop within adm using national agricultural statistics")
```
\normalsize

\pagebreak
```{r fig_area_share}
fig_area_share
```

\pagebreak
```{r fig_yld_outlier}
fig_yld_outlier
```

\pagebreak
```{r fig_yld_winsor}
fig_yld_winsor
```

\pagebreak
```{r fig_yld_winsor}
fig_yld_winsor_2000
```



\elandscape



# References


