---
title: "Global-to-local GLOBIOM procedure"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: ../Common/word_styles_01.docx
    
---

```{r setup, include=FALSE}
library(rprojroot)
root <- find_root(is_rstudio_project)

library(knitr)
knitr::opts_chunk$set(
  fig.width=12, fig.height=8,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"FigTabMap/out-"),
  dev = "CairoPNG",
  dev.args = list(CairoPNG = list(bg = "transparent"))
  )

library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

#source(file.path(root, "Code/Fig.R"))
#source(file.path(root, "Code/Tab.R"))
source(file.path(root, "Code/Map.R"))
```

# Introduction
GLOBIOM has been designed as global asssessment tool. Most input data is defined at the [0.5 degree] spatial level. Recently, there has been increasing demand for regional and subnational assessments that focus an a specific geographical region. Examples are land use change studies for Brazil, Congo, Ethiopia and Indonesia and an assessment of the water-land-energy nexus in the Zambezi and Indus river basins.The global datasets that are normally used for assessments are too coarse to arrive at the level of detail that is demanded for this type of studies and therefore needs to be replaced more detailed information that have a higher spatial resolution, if available.  

This document describes the various steps to prepare GLOBIOM for regional or subnational assessments. Refining the land use and land cover data for subnational assessments can be very time and resource-demanding and demands a certain craftsmanship. It is recommended to only follow the full process for relative large projects that take at least around two years. It is also advisable to involve local partners and stakeholders to provide input data (e.g. land use maps and production statistics) and validate processed information (e.g. updated land use map that is used in GLOBIOM) 

The use of national, country, subnational, region, district, zone, etc can be somewhat confusing in this type of analysis. For practical purposes and as most data is country specific, we will refer to 'national' level analysis when we mean countries or a croup of countries for which the analysis is conducted (e.g. South Asia). We will use the term administrative zones, when we refer to districts, provinces or zones that form divisions of a country (for practical reasons, we disregard that the administrave zone level 0 is used to indentify countries). 


# Data sources
Although the sources of data may vary, the following three pieces of data are needed to optimally improve the input structure of GLOBIOM for Global-to-local analysis:

## Land cover maps
The global land cover map that is used by GLOBIOM is very coarse and is not detailed enough for subnational analysis. It is therefore very important to try to find high-resolution land cover maps. Land cover (and land use) maps are often prepared by specialist government agencies (often under the Ministry of Agriculture or Forestry) or companies. The latter are sometimes contracted by international institutions or donors to provide new land cover maps [@Haack2015]. Sources for recent land cover maps are the Regional Centre of Mapping of Resources for Development (RCMRD) geoportal (http://geoportal.rcmrd.org.and) and Servir Global (https://www.servirglobal.net). 

Recently, the European Space Agency (ESA) has released a new time series of consistent global LC maps at 300 m spatial resolution on an annual basis from 1992 to 2015 (https://www.esa-landcover-cci.org/), which can also be used if national maps are not available. More research is needed to determine whether it is better to use one source of consistent data, such as the ESA for all case studies, albeit with (slightly) higher resolution, or use national data that is probably more precise. 


## Agricultural statistics
Agricultural statistics of developing countries are often not easy to find and demands a pragmatic approach. It works best to consult and combine several sources, including:

- Websites of national statistical agencies.
- Websites of the Ministry of agriculture, environment, fishery, etc.
- Websites of national agricultural research institutes.
- Agro-maps: http://kids.fao.org/agromaps/. An FAO initiative to collect subnational data on harvested area, production and yield. Contains lots of information although, quality and sources are not always clear. 
- Country stat: https://www.countrystat.org/. Another website of the FAO with subnational agricultural statistics. Contains much less information than Agro-maps. Not clear if the data is the same or different.
- Contacting CGIAR institutes that are active in the region.
- Contacting any of the national organisations listed above, preferably by means of a personal visit.
- Large companies that are active in the agricultural sector (mainly for statistics on cash crops such as palm oil, tea and coffee)


## Household surveys
In particularly useful as a source of information are household surveys that contain information for a large number of units, which are in most cases households. However, in most recent surveys, it is not uncommon to also provide data at the level of household members, plots or fields and communities as well as data from additional sources such as geo-spatial information that can be linked using geo-codes. Nationally representative surveys, preferably stratified at the district or zonal level, are the most useful source of information because they provide information for the complete country or region, which can be mapped to the simu level in GLOBIOM.

In particular three initiatives present nationally representative household information that can be used as a source of information. These are described below. Some of these datasets only cover a handfull of countries, while others have a wider coverage. A number of other data collection initiatives (e.g. IMPACT-lite; ILRI/CCAFS-CGIAR; RIGA, FAO-World Bank) might also present useful and interesting information but are not first choice because they address only a limited number of countries, cannot be compared across countries or are not nationally representative.

### The Living Standards Measurement Study (LSMS)
The Living Standards Measurement Study (LSMS) is an ongoing household survey program supported by the World Bank. Since the 1980s the LSMS surveys have been organised in a large number of countries. Depending on the country and year, the surveys present household and/or regional information on production, consumption and prices.       

### The Living Standards Measurement Study - Integrated Surveys on Agriculture (LSMS-ISA)
The Living Standards Measurement Study - Integrated Surveys on Agriculture (LSMS-ISA) is a project funded by the Bill and Melinda Gates foundation and implemented by the World Bank to undertake nationally representative households surveys in a eight African countries: Burkina Faso, Ethiopia, Malawi, Mali, Niger, Nigeria, Tanzania and Uganda. In each country multiple survey rounds are supported that make it possible to analyse long-term developments and apply econometric panel approaches. 

The surveys cover a wide range of topics and are particularly designed to improve the understanding between agricultural, socio-economic status and non-farm income activities. They cover around 3000 to 5000 households per country and include questions on a large number of topics, including agricultural production (including fishery and livestock) at crop and plot level, diets and consumption patterns and household composition and income. The dataset also include the coordinates of each sampling site and can therefore be linked with and compared to geo-spatial datasets. The LSMS-ISA surveys already include an additional data file with geo-linked information on precipitation, temperature and infrastructure.

### The Demographic and Health Surveys (DHS)
The Demographic and Health Surveys (DHS) Program is a large survey program to collect nationally representative information on fertility,  family  planning,  maternal  and  child  health,  nutrition, mortality,  environmental  health,  HIV/AIDS, malaria, and provision of health services funded by USAID. To date, the program has collected data for more than 90 countries in more than 300 surveys. The advantage of the DHS surveys is that they can easily be compared internationally because of the use of (nearly) similar questionnaires. A disadvantage is the limited information on agricultural activities, income and infrastructure. Recent DHS data has been geo-coded and can therefore be easily depicted on a map (see example below). Several recent studies have applied spatial interpolation techniques to provide higher resolution maps. 


# Folder structure
As we are using data from a variety of sources, it is important to apply a coherent and standardised structure for the location of the data and related R code that are used to process the information. We use a fixed structure for each of the target regions, which in most cases willl be a country. For practical reasons it is useful to use different main folders for data and scripts.

As most data refers to a country, we consistently use the ISO 3166-1 alpha-3 codes to indicate the region (https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3). This means that 'iso3c' in the text below should be replaced with the respective country code.The R package countrycode  (https://cran.r-project.org/web/packages/countrycode/index.html) can be used to transer iso3c codes in any other use country identifiers (e.g. ISO 3166-1 numeric, UN and FAO).

## Data folder
two main folders, one for the dataFor each, target region, we have two main folders thdata and code are organised stored in four main foldersusing the same folder names but are stored in two different main folders. The folder with all the da


## Code folder

## Documentation of data sources and procedures
For reproducability, a document with the title "Data_processing_sources_and_notes_iso3c.rmd" is created that describes important assumptions, gaps and procedures that are associated with the data preparation. It also lists all data sources that are used. Each document is accompanied by a *.bib file, created with Mendeley, that contains all the detailed information of the data sources.   

# Data processing pipeline
Figure X shows a flowchart for all the steps to prepare the data. We will discuss them one by one. 

 number of parallel processes. 


# GLOBIOM set up

# Regional module
- Creates one GDX file with all information for the region you are studying. 
- Add some lines in the aggregation phase, where regional data is replaced by new updated information.
-30 world regions

## Folder structure
Add a country case study folder in GAMS that will be used for all (sub)national statistics. It will have the same structure as a standard global data folder in GLOBIOM.

## Trade
Have to decide if you would keep one region or split countries. In latter case, adopt regional trade block and model trade with total block not bilaterally.
 
## Consumption


## FROM ALINE's procedure it appears the following data is needed
- Shares of Simus into adm units => calculate by overlay of GADM/GAUL and simu polygon
- Share of Simus in land cover class (already calculated)
- Area by adm
- Production adm
- Animals (dairy and milk) by adm (heads)
- Milk and meat by adm
- Statistics by adm (prod, area, animal count)
- Yield potential (ss) frop EPIC
- Suitability map for certain crops
- Transportation costs





# STEP 1: Collect (sub)national data
Preferably the following information should be updated in GLOBIOM:
- Area and production for each product (see Annex) by administrative unit in 2000 (and 2010)
- Land cover map classes by simulation unit for 2000 and (2010) [@Are there standard classes?]
- Protected areas map.
- Costs. At the moment only fertilizer and irrigation costs. Additional household data might improve the present information that is used in GLOBIOM, which is provided at a very high level (national/regional.)
- Productivity [?]

It highly depends on the region, which data is available and at which level. We recommend to take a pragmatic approach, using data from household surveys (e.g. LSMS-ISA, LSMS, etc), national statistics and specialised ministries and institutes. The following sections describe some rules of thumb that can be used to update the information.

The abbreviations (as, am, hs and lc) are used in the names of the mappings and files to refer to the source.

## Agricultural statistics (as)

## Agro-maps (am)

## Household surveys (hs)
## Spatial data (sd)



National statistics and household surveys provide the best information for detailed information on agricultural production. In most cases it should be possible to collect information on total production, crop area and/or harvested area at the second administrative level (e.g. province). Sometimes it will be necessary to add additional crops that are not part of the standard GLOBIOM crop classification (e.g. teff in Ethiopia) 

When production data is not available, another approach can be to collect information on average consumption or diets and combine this with population density maps to approximate total consumption at the level of the administrative unit. As in many poor countries, the agricultural sector is characterised by subsistince farming, consumption is a reasonable proxy for production in case no other information is available.




## Land cover map
Many countries publish detailed land cover maps, that provide information on the location of crop land, forest land, agro-industrial plantation (consessions), timeber plantations (concesstions), forest land, urbanised land and unused areas (i.e. rock). These maps can be used to improve the land cover map in GLOBIOM at the simulation unit level. 


# STEP 2: Select administrative regions and download map
## Introduction
To improve the modelling with GLOBIOM, we want to use detailed information that is available at the subnational level. Data can defined at the grid level with resolutions of 5 arcmin or 0.5 degree, or at some the administrative area. Administrative areas include the country level and lower level subdivisions such as provinces, districts and departments ofting using country specific names. National statistics (and sometimes international institutions, @See section X) often provide information on agricultural production such as output, yield and inputs at an administrative area lower than country level. Alternatively, as will be shown in section @X, the researcher can also calculate these aggregates if detailed household surveys are available.

The size of administrative areas differ by country. For example, the first administrative area in Ethiopia are nine regional states (or *Kililoch*) that are based on ethnic territoriality (https://en.wikipedia.org/wiki/Subdivisions_of_Ethiopia). These are fairly large areas. In contrast, in Tanzania, the first administrative area is made up of thirty regions (or *mkoa*), which are much smaller in (https://en.wikipedia.org/wiki/Subdivisions_of_Tanzania). To select the appropriate region, the researcher needs to study administrative division of the target country and compare this with available statistics. It is also important to check the extent to which adms have changed over time.

```{r Fig_1, fig.show = 'hold', out.width='33%'}
# https://bookdown.org/yihui/bookdown/figures.html for figures side by side
ETH_adm
TZA_adm
```

After the choice is made, the appropriate map(s) can be downloaded. There are two sources for adm maps: the GADM database of Global Administrative Areas (http://www.gadm.org) and the Global Administrative Unit Layers (GAUL) prepared by the FAO (http://www.fao.org/geonetwork/srv/en/metadata.show?id=12691). We prefer to use the GAUL because previous GLOBIOM input maps, in particular the simu information is prepared using GAUL. Furthermore, GAUL is available at an annual basis and therefore takes into account the changes in adms over time, while GADM tends to present only the most recent information. The latter is important as we need adm information for the base year of GLOBIOM (2000) and later years to validate the simulation (e.g 2010). 


## Script
__Location: Code/iso3c/Maps/__
__Select_GAUL_iso3c.r__. This script extracts the country specific map from the GAUL. Different levels of administrative boundaries can be selected and the maps will be automatically saved in user-selected folder for later use.


# STEP 3: Create mappings to harmonise data
## Introduction
Although often from the same data source (e.g. national statistical agencies), definitions of administrative areas, agricultural crops and inputs (e.g. types of fertilizer) are no consistent across household surveys, agriculatural statistics and GADM. In order to harmonise the data, it is essential to create mappings that link information from various sources. For further use all mapping are gathered in a single Excel file (Mappings_iso3c.xlsx), where each sheet contains a different mapping. Below, each mapping is briefly discussed. Note that often mappings will only be created when you are processing and analysing the data in STEP @X and @X.

We use the following conventions to create the mappings:

- The abbreviations for the three main sources are used in the mappings to indicate the source of the data: agricultural statistics (ag), household surveys (hs) and land cover maps (lc)
- All names from the three sources are capitalised to remove the problem that sometimes parts of region and crop names are capitalised in one source, while they are not in others (e.g. Mzuzu city versus Mzuzu City).
- 
- Lists of GLOBIOM crops and inputs are the same as in GLOBIOM and therefore only partially capitalised.

### Land cover classes: MWI2GLOBIOM_lc

### Administrative areas: iso3c2ADM_2000
In order to link data from the land cover maps, agricultural statistics and household surveys, it is important to create a list with administrative areas that will be used throughout the analysis to link data. For mapping purposes it is most convenient to use the GAUL area names (and code) as a basis. As names are often spelled differently, it is essential to prepare a concordance table that links all versions of adm names that are used by the various data sources. The table is stored in a sheet called __iso3c2ADM_2000__ in the __Mappings_iso3c.xlsx__ file. 

### Crops: MWI_as2FCL
Map all crops that are produced in the country to FAOSTAT crops. FAOSTAT provides a commodity list (FCL) that classifies crops, livestock and processed products (http://www.fao.org/economic/ess/ess-standards/commodity/en/),   which can be aggregated to GLOBIOM crop and livestock categories. Mapping country specific items to the FAO classification is sometimes not straigtforward and requires some thinking and creativity. Many tropical countries produce a large variety of beans (e.g. ordinary beans, ground beans, etc), which are all mapped to Beans, dry (BeaD). Similarly, cow peas and pigeon peas are typically mapped to chick peas [@check].

###

# STEP 4: Downscale land cover maps to SIMUs 

## Introduction
Information on land cover are essential for modelling land use change with GLOBIOM. The standard land cover map in GLOBIOM is produced by @ADD and has a resolution of @ADD. Although useful for global assessments, the spatial resolution is too coarse and imprecise to analyse land use change at the subnational level. For this reason, it is important too replace the land cover map for the target country with updated and country specific information that is often published by national agencies and international institutions such as the FAO and the World Bank. 

To use the new land cover maps in GLOBIOM it is important to (1) the harmonise land cover classification and (2) map the data to the GLOBIOM SIMUs. Land cover information in GLOBIOM is provided at the six classes that are also used by the IPCC for reporting [@check] Table @X presents the six classes with definitions. This is a standard classification that in most cases will also be used for country specific land cover maps. It is also possible that a more detailed classification consisting of  12–15 country-specific classes is used. In such cases, the researcher needs to aggregate the data first before it can be processed further.[^1] The next step is to aggregate the land cover information to SIMU level and write them to GDX.

## Script
__Process_land_cover_map.r__. This script loads a country specific land use map, aggregates it to SIMU level and saves the information in GDX format. For the script to work, the land cover data needs to be presented in raster format. The following procedure is used to aggregate the land cover data:

1. Load the global SIMU 5 armin polygon map and select the target country using the iso3c country number. Load the country specific GADM map (only needed for making maps).
2. If needed, reproject the the SIMU polygon and the GADM map to the target country projection.
3. Overlay the land cover map with the SIMU polygon map and calculate the share of each land cover class per SIMU. We assume that all raster cells have the same size (which is approximately true if the area is not too large and area is location at a distance from the poles) and therefore we can use the relative number of cells by land cover class as a measure of relative land cover area.
4. Save the data in GDX format.


__Table X: Standard Land cover classes__

```{r Tab_1}
Tab_LC <- read_csv(file.path(root, "Data/Standard_land_cover_classes.csv"))
pander(Tab_LC)
```
###### Source: @Haack2015


```{r, message = FALSE, results='asis'}
getwd()
library(readxl)
library(knitr)
library(xtable)
products <- read_excel(file.path(root ,"Data/GLOBIOM_products.xlsx"))
kable(products)

```


# STEP 5: Processing agricultural statistics
## Introduction
Often data from national statistical agencies, in particularly historical information, is only available in a format that is difficult to process, mainly pdf, diagrams or tables on a website. Fortunately, there exists a number of tools that make it possible to 'scrape' the information so there is no need to manually copy the data.

- Tabulizer is an R package that makes it possible to load a pdf into memory and extract tables from prespecified pages. This requires a bit of coding and requires high quality pdf documents (i.e. not scanned): https://ropensci.org/tutorials/tabulizer_tutorial.html
- There are a number of R (and Python) packages that make it possible to scrape tables from websites, most relevant are Rselenium and rvest. Using these packages does require some experience with R but documentation can be found on the internet.
- There are a number of software solutions to digitize graphs that work pretty well. A good example is WebPlotDigitizer (http://arohatgi.info/WebPlotDigitizer)

There is not one recipe or method on how to process, harmonise and clear the agricultural statistics data but the following steps probably have to be taken in any case:

- Scrape, download and digitize the data using any of the tools above to get a number of files with raw data.
- Reformat the data. For easy processing it is highly recomendable to convert all data files into the 'long' format where each variable (e.g. year, item, region and value) is stored in separate columns instead of using the 'wide' or 'matrix' format, where one variable is mapped horizontally. 
- Harmonise the data by applying the mappings from STEP [@X]. Often this involves aggregation over subsectors for which GLOBIOM uses a higher level aggregate.

## Scripts
There are a number of packages to perform the tasks listed above. In particular a set of packages referred to as 'tidyverse' have specifically designed for reformatting and 'slicing' data [@REF]. The [@XX] scripts contains some examples. 


# STEP 5: Analysing and cleaning agricultural statistics
## Introduction
After processing, it is important to analyse, and if needed, clean the agricultural statistics. We use national figures from FAOSTAT to benchmark the regional statistics on crop production, area, yield and heads of livestock. Since data from FAOSTAT represent national totals and averages, they should be comparable to the regional information. In case there are large differences, a decision has to be taken on how to deal with the data. As GLOBIOM relies heavily on FAOSTAT data, it will be leading in most cases. It is also important to check the data for outliers. Extremely high yields in one district that perhaps are caused by data errors can have dramatic consequences for the allocation of land. Finally, plotting and mapping the data also helps to get a 'feeling' about the target region as it reveals the relative importance of crops and the mayor crop areas in the region.

We conduct the following analysis and cleaning

- Importance of crops at national level. We calculate the share of crop area in total national crop area using agricultural statistics and FAOSTAT. This will give an impression which crops are most important to combine with the land cover maps. 
- Importance of regions by crop. For each crop we calculate the crop area per adm. This will give information which regions produce the most of a crop. It will, for example, reveal the 'bread basket' of a country. We also prepare maps for each crop. 
- Importance of crops within adms. We calculate the share of crop area of total adm crop area to determine if there is specialisation within regions.
- Distribution of yield for each crop in comparison to FAOSTAT. Boxplots diagrams reveal if yield of adms can be regarded as outliers. Comparison with FAOSTAT shows if the adm distribution is comparable to the national average as presented in FAOSTAT. We winsor all yield to the crop-specific IQC.
- Comparison of totals with FAOSTAT. To investigate if the regional information from agricultural statistics is comparable to FAOSTAT we compare national level production, area, yield and heads from FAOSTAT with totals from the agricultural statistics.



# STEP 6: Processing household surveys
##Introduction






# STEP X: Create cost database

# STEP X: Consistenty checks
- Compare ag land cover with ag production.
- Compare population density with land cover and ag production
- Compute share of 


# STEP X: Downscale land use and production data to Simu
## Introduction
Code in GAMS is provided to downscale and harmonise the land use and land cover data to the GLOBIOM simu levels. Downscaling is performed using priors that provide information on the suitability of certain crops or transportation costs. Downscaling is mostly down for crops as only limited information tends to be available on livestock. It is a separate optimization problem for each of the administrative units for which detailed production data is collected. 

The optimization is made up of the following steps:

1. 

To allocate the cropland area from the national agricultural statistics a 





# ANNEX
- GLOBIOM Crops
- GLOBIOM Inputs
- List of key files (Mappings_iso3c.xlsx)
- List of scripts and purpose


# Footnotes
[^1] This can easily be done with R using the raster package. At the moment no code is provided for this. 