---
title: "Global-to-local GLOBIOM procedure"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
#output: pdf_document
output:
  word_document:
    reference_docx: ../Common/word_styles_01.docx
bibliography: ../common/Global-to-local-GLOBIOM-MWI.bib    
---

```{r setup, include=FALSE}
library(pacman)
p_load(rprojroot)
root <- find_root(is_rstudio_project)

p_load(knitr)
knitr::opts_chunk$set(
  fig.width=16, fig.height=12,
  dpi = 300,
  echo=FALSE, warning=FALSE, message=FALSE,
  fig.path = file.path(root,"FigTabMap/out-"))

p_load(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

#source(file.path(root, "Code/Fig.R"))
#source(file.path(root, "Code/Tab.R"))
source(file.path(root, "Code/Paper/Map.R"))
```

# Introduction [TO BE EXPANDED]
Information on the location and size of crop area is essential to assess future agricultural growth and food security. Unfortunately, for most countries and in particular for developing countries, detailed agricultural statistics, such as crop-specific area, output and yield are often not available or only presented at highly aggregate levels (e.g country, district or province). Only very rarely is such data presented in map form with the high resolution that is required for land use modelling and scenario analysis. 

Recently, there have been several initiatives to produce global and spatially explicit land use maps for a large number of crops [@Monfreda2008, @Portmann2010, @You2014a]. Most these products use spatial information on land cover, suitability and irrigation to downscale (sub)national agricultural statistics to the grid level. A comparison revealed that  crop-specific harvested area and yield can differ considerably and mostly result from variation in input data, in particular crop masks and downscaling methodologies. [@Anderson2015]. 

A major advantage of the abovementioned datasets is that they have global coverage and therefore are often used by global integrated assessment models as base year input to model future land use change [@ADD]. On the other hand, due to their broad coverage, the global downscaling approaches mainly rely on global land cover masks and are restricted in their use of country specific data and expert input.[^1] Although most inititives use subnational statistics,  none of them incorporates other types of country specific agricultural information that improves the downscaling process. Particularly interesting  are high resolution land cover maps based on sattelite information and (nationally representative) household and farm surveys that contain geo-coded information on the location of crops production. The latter are increasingly available for a number of African countries [@Carletto2015].

For this reason, global land use maps are less useful or can serve, at best, as a starting point for national land use assessements. These studies are more focussed and often involve a proces of participatory scenario development in which stakeholders are asked to validate input data [@Rutten2014, @PALALZZO]. At this stage, local experts frequently point out mistakes and provide updated information such the location of plantations, updated land cover and land use maps and the location of irrigated areas [@CONGO PAPER]. To further improve the land use model a flexible procedure is needed that is able to incorporate all available national information. 

In this paper, we present such a model using Malawi as a case-study. With on average X USD per capita, Malawi is one of the poorest countries in the world. It's economy heavbily depends on agriculture, which makes up  one-third of gross domestic product (GDP) and two-thirds of employment [@CHECK]). The main food crop is maize, which is cultivated by the majority of small-scale farmers throughout the country. [@Tobacco]. With X inhabitants per square kilometer it is the X most densily populated country in Africa. With a projected increase in population size from X in X to X in 2050, the pressure for land will become increasingly larger. A better understanding on present land use can support land use change analysis. 

Our model builds on the Spatial Production Allocation Model (SPAM) presented in You et al. [-@You2006; -@You2009; -@You2014a], which uses  a cross entropy framework to allocate subnational land use information to 5 arcmin grid cells. It distinguises between four different crop systems: high-input irrigated, high-input rainfed, low-input rainfed and subsitence rainfed, each is characterised by different input output combinations. We extent the model by incorporating spatial explicit information on crop area from sattelite imagery and new irrigation maps based on irrigation surveys. Further, similar to @You2006, we use additional statistics (in our case from household survey) to validate the results from the allocation procedure. In a next step, we incorporate this data to improve the model [@check if possible].  


The structure of the paper is as follows:


[^1]: An exception is SPAM, which incorporates the feedback from experts on preliminary results into the model.  


# The spatial production allocation model

Our spatial allocation model refines the most recent version of SPAM [@You2014a, @Wood-Sichra2016]. SPAM uses a cross entropy objective function that expresses land allocation shares as probabilities, This function is minimized subject to a set of constraints that capture the available information on location of certain crops and crop systems (e.g. crop cover and irrigation maps) or the likelihood a crop is grown in a certain area (e.g. suitability maps and subnational land use statistics). Although previous work uses country specific data sources for subnational land use statistics, they mainly rely on global sources to define the land allocation constraints. In this study, we were able to collect more detailed information, which allowed us to refine the allocation procedure. In particular, we collected a high resolution land cover map, which reveals the location of four crop groups and created new crop specific irrigation maps for Malawi. [@PERHAPS UPDATE OF SUITABILITY MAPS]. In the sections below we present the adjusted SPAM model as well as the calculation of priors, which are needed to start the cross entropy optimization. To remain comparability with @You2014a, we use the same variable names and indices.

5 x 5 arcmin (around 9 x 9 km and around 8500 hectares when measured at the equator)


## Cross entropy framework and constraints
For the cross entropy framework can be implemented, the allocated area needs to be converted into a probability with value between 0 and 1. This is done as follows:

$$ s_{ijl} = \frac{A_{ijl}} {CropArea_{jl}}$$
where, $s_{ijl}$ is the area share allocated to grid cell identifier $i = 1,2,3,...$ and crop identifier $j = maize, cassave, rice,...$ at input level $l=irrigated, rainfed-high input, subsistence$. $A_{ijl}$ is the area allocated to grid cell $i$ for crop $j$ at input level $l$. $CropArea_{jl}$ is the total physical crop area in Malawi for crop $j$ at input level $l$.

The spatial allocation model can then be defined as a non-linear optimization problem in which the error between prior information on allocated area shares ({$\pi_{ij}$) and allocated area shares ($s_{ijl}$) is minimized, subject to a set of constraints. The objective function is defined as follows:

$$\min_{s_{ijl}} \quad CE(s_{ijl}, \pi_{ijl})= \sum_{i} \sum_{j} \sum_{l}s_{ijl} \ lns_{ijl}- \sum_{i} \sum_{j} \sum_{l}s_{ijl} \ ln\pi_{ijl}  $$
subject to:

(i) A constraint defining the range of permitted physical area shares:

$$0 \leq s_{ijl} \leq 1$$

(ii) A constraint (@X), which ensures  ensures that the sum of allocated physical area shares within grid cells sum to one. This ensures all physical crop area is allocated to grid cells. 

$$\sum_{i}s_{ij}=1 \qquad \forall j \quad \forall l$$

(iii) A constraint, which specifies that the sum of allocated physical area for each crop group, $c=rice, sugar \, cane,...,$ and production systems is lower or equal than the actual crop area ($Avail_{ij}$) that is available for each crop group $c$ in grid cell $i$. This equation extends the one in @You, by adding spatialy explicit information on the location of specific crops in the country.    


$$\sum_{j \in c} \sum_{l}CropArea_{jl} \times s_{ijl} \leq Avail_{ic} \qquad \forall i$$

(iv) A constraint [@ADD suitability equation]


$$CropArea_{jl} \times s_{ijl} \leq SuitArea_{ijl} \qquad \forall i \quad \forall j \quad\forall l$$

(v) A constraint, which ensures that the sum of allocated physical area over all production systems within a  sub-national geopolitical unit,$k=1,2,3,...$ (in this case administrative zone 2 level) is equal to the  physical area information ($SubCropArea_{jk}$) for commodities $J$ for which information is available from sub-national land use statistics. Crops for which only national level information is available are not affected by this constraint. 


$$\sum_{i\in k} \sum_{l}CropArea_{jl} \times s_{ijl} = SubCropArea_{jk} \qquad \forall k \quad \forall j \in J$$

(vi) A constraint, which allocates spatially explicit information on irrigated areas ($IRRArea_{ijl}$) for  irrigated crops ($J$) that belong to the irrigated crop system in grid cells $I$. In contrast to $You2014, who do not have crop specific irrigation area, we have detailed location and crop specific data on irrigated areas, which we incorporate in the spatial allocation model. 

$$CropArea_{jl} \times s_{ijl} = IRRArea_{ijl} \qquad \forall i \in I \quad \forall l = irrigated \quad \forall j \in N$$


## Calculation of priors
[TO ADD. At the moment I use population density but for the high-intensity we can follow the approach of @You2014a]


# Data
We combined information from a number of different sources to run the model described above (see Table X for an overview). A key input for the analysis is a high resolution land cover map that provides detailed information on the location of agricultural production. Recently, several new land use and land cover maps were produced for Malawi that cover the period 1990-2010 [see @Haack2015 for a comparison]. We decided to use the maps produced by the FAO [-@FAO2013] that is available for 2010. The advantage of this map in comparison to others is that it has a very high resolution of 30x30 meter and therefore provides spatially explicit information on the location of four crop groups (rice, sugar cane, tea & coffee and other crops) in Malawi. We aggregated the map to a resolution of @X and calculated the physical area share for each crop group per grid cell $i$.

Statistics on actual land use at the national level were taken from @FAO and extended with data from @Add, @add and @add for the administrative zone two level. We follow the approach of @Wood-Sichra2016 and aggregate all crop statistics to a set of standardised crops and subsequently allign the subnational data with national statistics from FAO. We were able to collect subnational data for XX out of the XX crops that are cultivated in Malawi. Annex X provides more information on the prepartion of the crop statistics.

Similar to @You2014a, we use @GAEZ suitability map, which can be mapped to the standardised crops (see @Wood-Sichra2016). [@or perhaps use IFPRI maps].

Finally, we used detailed information from the National irrigation master plan and investment framework [@Add], which presents a geo-referenced inventory of irrigation systems in Malawi, including the size of the irrigated area and crop, for the period @add. Where needed we compared and extended the inventory with data from @Add, which provides similar but less detailed information. As the irrigated production system in SPAM are defined as high-input system, we decided only to include the estate and large smallholder association estates and excluded small-scale farmer schemes. A comparison showed that the resulting area is comparable to @Siebert2013, which is used by @You2014a as input layer. To create the final irrigation map, we compared the FAO land cover map with the irrigation map and allocated the point data to grid cells with matching crops (i.e. sugar cane irrigated areas to sugar cane grid cells). Annex X provides more detailed information about the procedure. 


__Table 1: Data sources__

```{r Tab_data}

```


# Results

__Figure X: model results for selected crops__

```{r Map_crop_example}
plot_crop_raster_f(c("maiz", "rice"))

```

##### Grey area indicates land cover not suitable for crop production


# Validation
## Comparison with SPAM
[TO ADD]


## Comparison with LSMS-ISA


__Figure X: Comparison between land use allocation and LSMS-ISA__

```{r Map_validation}
plot_crop_val_f(c("maiz", "rice"))
```

## Updating of priors
[TBD]


# Discussion



# Conclusions
@Perhaps refer to limited crop and area statistics: Re-evaluating the Malawian Farm Input Subsidy Programme

# References


