---
title: "micro-GLOBIOM"
author: "Michiel van Dijk"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: ../Common//word_styles_01.docx
bibliography: ../common/micro-GLOBIOM.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The GLOBIOM team is facing an increasing demand for sub-regional (as opposed to global) assessments of deforestation, climate change and food security. Examples of ongoing work include a farm systems analysis for Ethiopia, deforestation scenarios for Brazil, Indonesia and Congo and modelling of consumption and diet change at the state-level for India. A new project along these lines is ISWEL, which aims to analyse the trade-offs between water, enery and food security (the nexus) in the Indus and Zambezi river basins. Sub-regional model studies and the often associated demand for socio-economic impact analysis (i.e. poverty, food security and health) can benefit from the use of plot, farm and households surveys that are increasingly available and provide micro-level information on consumption, production and income. 

Previously the use of this data was problematic because surveys only covered very small regions were organised in an ad hoc manner or were of low quality. Recently a number of efferts were undertaken to address these issues and collect nationally representative data using a systematic approach that allow for international comparisons and cover multiple years. Examples of such datasets are: the World bank Living Standards Measurement Study (LSMS), the related Measurement Study Integrated Surveys on Agriculture (LSMS-ISA) and the Demographic and Health Surveys (DHS). 

A second interesting and related source of information to improve both the sub-regional and global modelling is spatially explicit data on number of socio-economic indicators that are build using micro-level surveys. These maps can be either used as input into to the modelling process (e.g. population maps) or be contrasted with model outputs (e.g. climate hotspots with poverty and food security maps). Although, spatially explicit information on socio-economic indicators is relatively thin [@Otto2015], there is ongoing effort to improve data availability [@Azzarri2016].

The aim of this document is to present some first ideas on how to expand and deepen the modelling with GLOBIOM by integrating and linking micro-level and spatial information on socio-economic variables (for now  I will refer to this process as 'micro-GLOBIOM' for lack of a better description). The main idea is that such data can provide more detail on several key building blocks of the model, including the modelling of production and consumption. In addition, these data might also be used to construct a number of development related indicators that benefit from micro-level data, such as poverty and food security indicators. This document describes some of the possible directions we can take to use the available data inorder to enrich the modelling with GLOBIOM. It partially builds on on previous work done by colleagues that already explored this type of information, including farm systems modelling in Ethiopia, consumption modelling in India and forest modelling in Brazil, Congo and Indonesia as well as an additional literature survey. The document serves as a basis for discussion and is expected to be updated, refined and expanded as the thinking on this topics progresses. 

# Possible directions to enhance GLOBIOM with micro-data
Following the brainstorm discussion four, potentially interesting, approaches to develop 'micro-GLOBIOM' have been identified. 

## Combining GLOBIOM output with spatially explicit socio-economic indicators
As GLOBIOM presents spatially explicit results, it might be interesting to combine GLOBIOM output with spatially explicit socio-economic indicators. An example of such analysis could be the combination (overlay) of a map that depicts 'hotspots' of land use change and a poverty map to assess which people are most vulnerable. Unfortunately, spatially explicit socio-economic data is only available for a handfull of indicators. The main reason is that the required raw micro-level data to construct such maps is not available at high-resolution because it is too costly to collect by means of surveys. An interesting development in this regard is the collection of poverty data by means of mobile phone metadata [@Blumenstock2015] and sattelite imagery [@Jean2016], which will likely increase the availability of poverty maps at high resolution in the near future.  

### Potential data
For the following indicators spatially explicit socio-economic data is available for multiple countries or even at the global scale.

- Poverty: 
- Children stunted
- ChildrenConsumption?
- Population density: 


```{r maps}
# Add maps
```

### Examples of related work
Probably due to limited information, there are not many global modelling exercises that make use of spatially explicit socio-economic data. An interesting an recent example is Biewald2015 [-@Biewald2015], who analyse the impact of climate change on hunger for 2030. They use the MAgPIE model to provide spatially explicit projections on changes in the costs of food. These results are subsequently compared with a spatially explicit Vulnerability to Hunger Index under different scenarions. The index is constructed by combining maps on child underweight, child mortality and prevalence of stunding under five and projected into the future on the basis of national level poverty scenarios. 

# ADD info on TOA model

### Required investment
For the abovementioned indicators data can be easily downloaded and combined with GLOBIOM output. This can either being done within GAMS (I think) or by means of postprocessing in R or any GIS software (QGIS, ArcGis, etc).

## Update the supply side




## Improve demand side


## Produce socio-economic indicators






# Next steps

## Determine priorities 

# Spatial data sources


# Household level data sources
There are a number of data sources with plot, farm or household information that can be used to enhance the modelling of GLOBIOM, which are summarised below. All of the sources present information that is at least nationally representative and therefore can be used as input to country-level modelling. Regional coverage strongly differs across sources. Some datasets only cover a handfull of countries, while others have mode wider coverage.

## The Living Standards Measurement Study (LSMS)

## The Living Standards Measurement Study - Integrated Surveys on Agriculture (LSMS-ISA)
The Living Standards Measurement Study - Integrated Surveys on Agriculture (LSMS-ISA) is a project funded by the Bill and Melinda Gates foundation and implemented by the World Bank undertake nationally representative households surveys in a eight African countries: Burkina Faso, Ethiopia, Malawi, Mali, Niger, Nigeria, Tanzania and Uganda. In each country multiple survey rounds are supported that make it possible to analyse long-term developments and apply econometric panel approaches. 

The surveys cover a wide range of topics and are particularly designed to improve the understanding between agricultural, socio-economic status and non-farm income activities. They cover around 3000 to 5000 households per country and includes questions on a large number of topics, including agricultural production (including fishery and livestock) at crop and plot level, diets and consumption patterns and household composition and income. The data also includes the coordinates of the sampling sites and can therefore can be linked with and compared to geo-spatial datasets. The database already includes information several spatial variables such as precipitation, temperature and infrastructure.

# The Demographic and Health Surveys (DHS)
The Demographic and Health Surveys (DHS) Program is a large survey program to collect nationally representative information on fertility,  family  planning,  maternal  and  child  health,  nutrition, mortality,  environmental  health,  HIV/AIDS, malaria, and provision of health services funded by USAID. To date, the program has has collected data for more 90 countries in more than 300 surveys. Data is pr Recent data is  

DHS data are, however, weak on agricultural activities and income, as well as infrastructure. 

# Ad hoc surveys


# Thinking wider and broader: Global socio-economic modelling with GLOBIOM
There is an increasing awareness about the interaction and  of climate. More demand: LANCET, SUSFANS


Examples
GAEZ
GISMO
IMPACT - health modelling

Next steps
# References

