---
title: "Test report"
author: "Michiel van Dijk"
date: "8 November 2016"
output: 
  word_document:
    reference_docx: ../Common//word_styles_01.docx

# For more info:
# http://rmarkdown.rstudio.com/articles_docx.html
# https://scriptsandstatistics.wordpress.com/2015/12/18/rmarkdown-how-to-inserts-page-breaks-in-a-ms-word-document/
# https://scriptsandstatistics.wordpress.com/2016/08/29/how-to-number-and-reference-tables-and-figures-in-r-markdown-files/
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Heading 1
My report starts here.

# another heading 1

## Heading 2
And I have this next subsection

# another heading 1
## heading 2
### Heading 3

# Make a table
Create a subset of the mtcars data. 

and a code chunk to extract a subset of data from the mtcars data set.

 <br>

```{r code, message = FALSE}

library(dplyr)
# extract a subset of the mtcars data
df <- mtcars %>%
    mutate(make_model = row.names(mtcars)) %>%
    filter(cyl == 4) %>%    
    select(make_model, mpg, wt) %>%
    mutate(wt = wt*1000) %>%
    arrange(make_model) 

```

<br>

Add some text followed by a code chunk,
Then use *knitr::kable()* to create the table.

<br>
```{r table, echo = FALSE}
library(knitr)
kable(df) 
```

<br>

```{r cars}
summary(cars)
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
